---
title: <center><strong><h2>[MiR Training Session](https://twitter.com/miR_community)</h2></strong></center>
subtitle: <center><strong><h2>Getting to know ggplot2</h2></strong></center>
author: <center><strong><h1>[Shelmith Nyagathiri Kariuki](https://shelkariuki.netlify.app/)</h1></strong></center>
date: <center><strong><h2>July 03, 2020</h2></strong></center>
output:
  html_document:
    toc: yes
    toc_depth: '6'
---

Dataset: [Financial Inclusion in Africa](https://zindi.africa/competitions/financial-inclusion-in-africa)

### 0.1 Set up

```{r setup}

knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,warning=FALSE,
                      fig.width = 8, fig.height = 6)
```

### 0.2 Install and load the packages required

```{r, }

###  create a vector of packages to be installed
pkgs <- c("tidyverse","DT","lubridate","patchwork")

###  Check if there are packages you want to load, that are not already installed. 
miss_pkgs <- pkgs[!pkgs %in% installed.packages()[,1]] 

###  Installing the missing packages
if(length(miss_pkgs)>0){
  install.packages(miss_pkgs)
}

###  Loading all the packages
invisible(lapply(pkgs,library,character.only=TRUE))

###  Remove the objects that are no longer required
rm(miss_pkgs)
rm(pkgs)

```


```{r, message=FALSE, warning=FALSE}
# Set the theme
mir_theme <- theme(plot.title = element_text(size = 18,family = "Source Sans Pro Semibold", face = "italic", hjust = 0.5),
              axis.line = element_line(color = "black", size = 1),
              axis.title = element_text(size = 16),
              axis.text = element_text(size = 14),
              panel.background = element_rect(fill = NA),
              plot.caption = element_text(size = 14),
              legend.title = element_blank(),
              legend.position = "bottom")

```


### 1. Read in the dataset

```{r, message=FALSE, warning=FALSE}

df <- read_csv("Train_v2.csv")

```

### 2. Convert character variables to factor variables 

```{r, }

## 2.1. bank_account
df <- df %>% 
  mutate(bank_account = fct_relevel(bank_account, "No", "Yes"))

## 2.2. location_type
df <- df %>% 
  mutate(location_type = fct_relevel(location_type, "Rural" , "Urban"))

## 2.3. cellphone_access
df <- df %>% 
  mutate(cellphone_access = fct_relevel(cellphone_access,  "No", "Yes"))

## 2.4. gender_of_respondent
df <- df %>% 
  mutate(gender_of_respondent = fct_relevel(gender_of_respondent,  "Female" ,"Male"))


## 2.5. relationship_with_head
df <- df %>% 
  mutate(relationship_with_head = fct_relevel(relationship_with_head,  
        "Child","Spouse", "Parent", "Head of Household",  
        "Other relative",   "Other non-relatives" ))

## 2.6. marital_status
df <- df %>% 
  mutate(marital_status = fct_relevel(marital_status,  
        "Single/Never Married","Divorced/Seperated","Widowed",
        "Married/Living together","Dont know"))

## 2.7. education_level
df <- df %>% 
  mutate(education_level = fct_relevel(education_level,  
        "No formal education", "Primary education", "Secondary education", 
        "Tertiary education", "Vocational/Specialised training","Other/Dont know/RTA"))

## 2.8. job_type

```

### 3. Generate 2 new variables i.e monthly income and date they registered for xyz

```{r, }
 ## 3.1 monthly income

set.seed(2020)
income_values <- sample(c(5000 : 150000), nrow(df))

df <- df %>% 
  mutate(income = income_values,
         income = ifelse(job_type == "No Income", NA, income))

## 3.2 date of registering for xyz
date_vec <- sample(seq(as.Date('2014/01/01'), as.Date('2014/12/31'), by="day"), nrow(df), replace = T)

df <- df %>% 
  mutate(date = date_vec) 

## Generate the month
df <- df %>% 
  mutate(month = month(date, abbr = T, label = T)) 

```

### 4. Main graphs

#### 4.1 Bar graph

##### 4.1.1 Single categorical variable

Distribution of cellphone access

```{r, }

## First generate a table
tab1 <- df %>% 
  group_by(cellphone_access) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = round(n/sum(n) *100, 0))

## Plot a graph
plot1 <- ggplot(data = tab1, aes(x = cellphone_access, y = perc))+
  geom_bar(stat = "identity", fill = "maroon", width = 0.7)+
  geom_text(aes(label = perc), size = 5, hjust = 0.5, vjust = -0.25)+
  mir_theme +
  labs(title = "Distribution of cellphone access",
       x = "",y = "Percentage",
       caption = "Twitter:@Shel_Kariuki")+
  ylim(c(0,100))
plot1

#rstudio_blue <- "#4AA4DE"
```

Distribution of sample by country

```{r}
## table

tab2 <- df %>% 
  group_by(country) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = round(n / sum(n) *100,1))

## graph

plot2 <- ggplot(data = tab2, aes(x = country, y=perc)) +
            geom_bar(stat = "identity", fill = "maroon", width = 0.7) +
          geom_text(aes(label = perc), size = 4 , hjust = 0.5 , vjust = -0.25)+
          mir_theme +
        labs(title = "Distribution of sample by country",
             x = "", y = "Percentage",
             caption = "Twitter: @Shel_Kariuki")
plot2          

```



##### 4.1.2 Two categorical variables (main variable and grouping variable)

Distribution of bank account availability by country

```{r, }

## table

tab3 <- df %>% 
  group_by(country, bank_account) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(bank_account) %>% 
  mutate(perc = round(n / sum(n) *100,1))

## graph

plot3 <- ggplot(data = tab3, aes(x = country, y=perc, fill = bank_account)) +
            geom_bar(stat = "identity",  width = 0.7, position = "dodge") +
          geom_text(aes(label = perc), size = 4 , hjust = 0.5 , vjust = -0.25, 
                    position = position_dodge(width = 0.7))+
          mir_theme +
        scale_fill_manual(values = c("grey", "maroon"))+
        labs(title = "Distribution of bank account availability by country",
             x = "", y = "Percentage",
             caption = "Twitter: @Shel_Kariuki")+
   ylim(c(0, 50))
plot3   

```

#### 4.2 Line graph

##### 4.2.1 Single categorical variable

Distribution of registrations by month

```{r, , fig.width = 10}
## table
tab4 <- df %>% 
  group_by(month) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = round(n/sum(n)*100,1))

## graph
plot4 <- ggplot(data = tab4, aes(x = month, y = perc, group = 1)) +
          geom_point()+
          geom_line(stat = "identity", color = "maroon", size = 1)+
          geom_text(aes(label = perc), size = 5, hjust = 0.5, vjust = -0.25)+
          mir_theme +
          labs(title = "Distribution of registrations by month",
               x = "", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")+
  ylim(c(7.5,9))

plot4

```

Average age of respondents registering per month

```{r, }
 ## table
tab5 <- df %>% 
  group_by(month) %>% 
  summarise(avg_age = round(mean(age_of_respondent, na.rm = T),1))

## plot
plot5 <- ggplot(data = tab5, aes(x = month, y = avg_age, group = 1))+
          geom_point()+
          geom_line(stat = "identity", size = 1, color = "maroon", linetype = "solid")+
          geom_text(aes(label = avg_age), vjust = 0, hjust = 1)+
          mir_theme +
          labs(title = "Average age of respondents registering per month",
               x = "", y = "Average Age",
               caption = "Twitter: @Shel_Kariuki")+
  ylim(c(37.5,39.5))
plot5
```

##### 4.2.2 Two categorical variables (main variable and grouping variable)

Distribution of registrations per month and country


```{r, }
## table
tab6 <- df %>% 
  group_by(month, country) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(country) %>% 
  mutate(perc = round(n/sum(n)*100,1))

## plot
plot6 <- ggplot(data = tab6, aes(x = month, y = perc, group = country, color = country))+
          geom_point()+
          geom_line(stat = "identity", size = 1, linetype = "solid")+
          #geom_text(aes(label = perc), vjust = 0, hjust = 1)+
          mir_theme +
  scale_color_brewer(palette = "Spectral")+
          labs(title = "Distribution of registrations per month and country",
               x = "", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")

plot6

```


```{r, }
## table
# tab6 <- df %>% 
#   group_by(month, country) %>% 
#   count() %>% 
#   ungroup() %>% 
#   group_by(country) %>% 
#   mutate(perc = round(n/sum(n)*100,1))
# 
# ## graph
# plot6 <- ggplot(data = tab6, aes(x = month, y = perc, group = country, color = country))+
#           geom_point()+
#           geom_line(stat = "identity", size = 1)+
#           #geom_text(aes(label = perc), hjust = 1.5, vjust = 0)+
#           mir_theme+
#           scale_color_brewer(palette = "Spectral")+
#           labs(title = "Registrations by month",
#                x = "", y = "Percentage",
#                caption = "Twitter: @Shel_Kariuki")
# plot6

```

#### 4.3 Scatter plot

##### 4.3.1 2 continuous variables
 Distribution of Income and household size
 
```{r, }

plot7 <- ggplot(data = df, aes(x = household_size,  y = income))+
          geom_point(size = 1, color = "maroon", shape = 3)+
          mir_theme +
  labs(title = "Distribution of Income and household size",
       x = "Household Size", y = "Income",
       caption = "Twitter: @Shel_Kariuki")
plot7
```

##### 4.3.2 2 continuous variables and a grouping variable

```{r, }

plot8 <- ggplot(data = df, aes(x = household_size,  y = income, color = country))+
          geom_point(size = 1, shape = 3)+
          mir_theme +
          scale_color_manual(values = c("black","maroon","brown","yellow"))+
          labs(title = "Distribution of Income and household size",
            x = "Household Size", y = "Income",
          caption = "Twitter: @Shel_Kariuki")
  
plot8

```


### 5. Other neat tricks

#### 5.1 Reordering bar graphs by ascending order of the y axis

```{r, }
## table
tab2 <- df %>% 
  group_by(country) %>% 
  count() %>% 
  ungroup() %>% 
  mutate(perc = round( n/ sum(n)*100,0))
  
## graph

plot2b <- ggplot(data = tab2, aes(x = reorder(country,-perc), y=perc)) +
            geom_bar(stat = "identity", fill = "maroon", width = 0.7) +
          geom_text(aes(label = perc), size = 4 , hjust = 0.5 , vjust = -0.25)+
          mir_theme +
        labs(title = "Distribution of sample by country",
             x = "", y = "Percentage",
             caption = "Twitter: @Shel_Kariuki")+
        ylim(c(0,40))
plot2b 

```

#### 5.2 Flipping graphs

```{r, }
 ## graph
plot2c <- ggplot(data = tab2, aes(x = country, y = perc ))+
          geom_bar(stat = "identity", fill = "maroon")+
          geom_text(aes(label = perc), hjust = -0.5, vjust = 0.5, size = 4.5)+
          mir_theme+
          labs(title = "Distribution of sample by country",
               x = "", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")+
         coord_flip()+
         ylim(c(0,40))
plot2c

```

#### 5.3 Faceting plots

##### 5.3.1 Facet wrap: faceting a plot by one variable

Distribution of education level by country

```{r, , fig.width=14, fig.height=10}
## table
tab9 <- df %>% 
  group_by(education_level, country) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(country) %>% 
  mutate(perc = round(n/sum(n) *100,1))

##plot
plot9 <- ggplot(data = tab9, aes(x = education_level, y = perc))+
          geom_bar(stat = "identity", fill = "maroon")+
          geom_text(aes(label = perc), vjust = -0.25, hjust = 0.5)+
          mir_theme+
          theme(axis.text = element_text(size = 12))+
          labs(title = "Distribution of education level by country",
               x = "Education level", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")+
          facet_wrap(~country, ncol = 1)+
          ylim(c(0,70))
plot9          

```

##### 5.3.2 Facet grid: Faceting a plot by two variables

Distribution of gender by country and location type

```{r, , fig.width=10, fig.height=10}

## table
tab10 <- df %>% 
  group_by(location_type, country, gender_of_respondent) %>% 
  count() %>% 
  ungroup() %>% 
  group_by(country,location_type) %>% 
  mutate(perc = round(n/sum(n) *100,1))

##plot
plot10 <- ggplot(data = tab10, aes(x = gender_of_respondent, y = perc))+
          geom_bar(stat = "identity", fill = "maroon", width = 0.7)+
          geom_text(aes(label = perc), vjust = -0.25, hjust = 0.5)+
          mir_theme+
          theme(axis.text = element_text(size = 12))+
          labs(title = "Distribution of gender by country and location type",
               x = "Gender", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")+
          facet_grid(country ~ location_type)+
          ylim(c(0, 75))
plot10  
```

#### 5.4 Adding % marks on the y axis and the text labels, and altering the breaks of the y axis

```{r, }
## graph
plot2d <- ggplot(data = tab2, aes(x = country, y = perc ))+
          geom_bar(stat = "identity", fill = "maroon")+
          geom_text(aes(label = paste0(perc,"%")), hjust = 0.5, vjust = -0.25, size = 4.5)+
          theme(plot.title = element_text(size = 18, family = "Source Sans Pro Semibold",
                                          face = "italic", hjust = 0.5),
                panel.background = element_rect(fill = NA),
                axis.line = element_line(size = 1, colour = "black"),
                axis.text = element_text(size = 14),
                axis.title = element_text(size = 16),
                plot.caption = element_text(size = 14))+
          ylim(c(0,40))+
          scale_y_continuous(limits = c(0, 40), breaks = seq(0, 40, by = 5),
           labels = function(x) paste0(x, "%"))+
          labs(title = "Distribution of sample by country",
               x = "", y = "Percentage",
               caption = "Twitter: @Shel_Kariuki")
plot2d  
```

#### 5.5 Using patch work: a package for combining multiple plots

```{r, , fig.width=16, fig.height=10}
 
# p1 <-  plot2 / plot3
# p1
end <- (plot2 + plot3) / (plot5 + plot6)
end
```
